Homological Algebra\vspace{-1ex}

### 2019-06-18 Homological Algebra Notes/Hom Alg.tex
Update and create new file

@@

\begin{prop}
	Given $0\to A\to B\to C\to 0$ an SES, we get our long exact sequence in Tate cohomology.
\end{prop}
\begin{prop}
	When $G$ is cyclic, $\hat H^n(G,A)\cong\hat H^{n+2}(G,A)$
\end{prop}
\begin{prop}
	We have a nice cup product
	\[\hat H^m(G,A)\otimes \hat H^n(G,B)\to \hat H^{n+m}(G,A\otimes B)\]
\end{prop}
\begin{prop}
	For all $m$ there exists a ``shift module'' $S_m$ such that 
	\[\hat H^n(G,A)\cong\hat H^{n-m}(G,A\otimes S_m).\]
\end{prop}
\begin{thm}[Tate]
	Let $G$ be finite, $A$ a $G$ module such that 
	\begin{itemize}
		\item $\forall H\le G,\hat H^1(H,A)=0$
		\item $\forall H\le G, \hat H^2(H,A)$ is cyclic of order $|H|$
	\end{itemize}
	Then if $A$ generates $\hat H^2(G,A)$ then we get an isomorphism
	\[a\cup -:\hat H^n(G,\bbZ)\cong \hat H^{n+2}(G,A)\]
\end{thm}
Then setting $n=-2$, we get that 
\[\hat H^{-2}(G,\bbZ)\cong\hat H^0(G,A)=\coker N_G=A^G/NA\]
but then 
\[A^G/NA\cing \hat H^0(G,A)\cong H_2(G,\bbZ)=G^{ab}\]
where the last thing is the abelianization of $G$.

@@

@@

\newpage
\section{HW2 -- Due XXXXX}
\begin{prob}
	Let $R$ be a commutative ring, and $N$ be an $R$-module. Show that the following are equivalent:
	\begin{enumerate}
		\item $N$ is flat 
		\item $\Tor_i^R(M,N) = 0$ for any $i>0$ and any $R$-module $M$
		\item $\Tor_1^R(M,N) = 0$ for any $R$-module $M$
		\item $\Tor_1^R(R/I, N) = 0$ for any ideal $I \subset R$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Rightarrow$(b)$\Rightarrow$(c) is clear since $\Tor_i^R(-,N)$ is a left derived functor of $-\otimes N$. But when $N$ is flat, $-\otimes N$ is exact.

@@

\begin{prob}
	Let $(R, \frakm, k)$ be a commutative local ring and $M$ be a finitely generated $R$-module. Prove that the following are equivalent: 
	\begin{enumerate}
		\item $M$ is free
		\item $M$ is flat
		\item The map $\frakm \otimes_R M \to R \otimes_R M =M$ induced by the embedding $\frakm \subset R$ is injective
		\item $\Tor_1(k,M)=0$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Leftrightarrow$(b): First let $M$ be free (and finitely generated). Then $M\cong R^k$ for some finite $k$. 
	Then take an injection
\end{sol}

@@

\begin{prob}
	Let $k$ be a field of positive characteristic $p$. 
	\begin{enumerate} 
		\item Show that the group algebra $kG$ for a finite group $G$ is Frobenius (in fact, symmetric). 
		\item Show that the restricted enveloping algebra $\fraku(\gl_n)$ is Frobenius. 
	\end{enumerate} 
\end{prob}
\begin{sol}

>>>>

%Shade definitions
\theoremindent0cm
\theoremheaderfont{\normalfont\bfseries} 
\def\theoremframecommand{\colorbox[rgb]{1,0.9,.55}}
\newshadedtheorem{defn}[thm]{Definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% Customize Below %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{prop}
	Given $0\to A\to B\to C\to 0$ an SES, we get our long exact sequence in Tate cohomology.
\end{prop}
\begin{prop}
	When $G$ is cyclic, $\hat H^n(G,A)\cong\hat H^{n+2}(G,A)$
\end{prop}
\begin{prop}
	We have a nice cup product
	\[\hat H^m(G,A)\otimes \hat H^n(G,B)\to \hat H^{n+m}(G,A\otimes B)\]
\end{prop}
\begin{prop}
	For all $m$ there exists a ``shift module'' $S_m$ such that 
	\[\hat H^n(G,A)\cong\hat H^{n-m}(G,A\otimes S_m).\]
\end{prop}
\begin{thm}[Tate]
	Let $G$ be finite, $A$ a $G$ module such that 
	\begin{itemize}
		\item $\forall H\le G,\hat H^1(H,A)=0$
		\item $\forall H\le G, \hat H^2(H,A)$ is cyclic of order $|H|$
	\end{itemize}
	Then if $A$ generates $\hat H^2(G,A)$ then we get an isomorphism
	\[a\cup -:\hat H^n(G,\bbZ)\cong \hat H^{n+2}(G,A)\]
\end{thm}
Then setting $n=-2$, we get that 
\[\hat H^{-2}(G,\bbZ)\cong\hat H^0(G,A)=\coker N_G=A^G/NA\]
but then 
\[A^G/NA\cong \hat H^0(G,A)\cong H_2(G,\bbZ)=G^{ab}\]
where the last thing is the abelianization of $G$.

@@

Bah I missed something here.

\section{June 7, 2019}
Today Christine and Tejas are going to talk about \textit{Abelian Sheaf Cohomology}.

\subsection{Sheaves}
\begin{ex}
	The idea here is that we are generalizing functions on spaces. For instance, consider $X=S^2$, the 2 sphere.
	Then take our fuctions to be ones satisfying $f\in\bbR[x,y,z]/(x^2+y^2+z^2-1)$ and open sets to be the usal ones on the spheres.
\end{ex}
\begin{defn}
	A \textbf{presheaf} of ableiang groups on a toplological space $X$ is a functor 
	\[\calF:\mathbf{Ob}(X)^{op}\to \Ab\]
	where $\mathbf{Op}(X)$ is the poset category on the open sets of $X$.

	Generally we suppres the space and the fact we are using abelian groups.
\end{defn}
\begin{defn}\label{defn-sheaf}
	A \textbf{sheaf} of abelian groups on $X$ is a presheaf such that the following are satisfied:
	\begin{itemize}
		\item (\textbf{Locality}) If $\{U_i\}$ is an open cover of an open set $U$ and if $s,t\in\calF(U)$ such that 
		\[s|_{U_i}=t|_{U_i},\forall i\]
		then $s=t$.
		\item (\textbf{Gluing}) If $\{U_i\}$ is an open cover of $U$ and if for each $i$ there is a $s_i\in\calF(U_i)$
		such that $s_i|_{U_i\cap U_j}=s_j|_{U_i\cal U_j}$, then there is a \textbf{global section} $s\in \calF(U)$ such taht $s|_{U_i}=s_i$.
	\end{itemize}
\end{defn}
If we like categories more (WE DO!) then we can equivalently define it thusly:
\begin{defn}\label{defn-sheaf-2}
	A sheaf is a presheaf $\calF$ such that for any open $U$ with cover $\{U_i\}$,
	\[\calF(U)\to \prod_i\calF(U_i)\stackrel{\to}{\to}\prod_{i,j}\calF(U_i\cap U_j)\]
	where, more generally, we could write this using $U_i\times_UU_j=U_i\cap U_j.$
\end{defn}

What is a morphism of sheaves, then?
\begin{defn}
	Let $\calF$ and $\calG$ be sheaves on $X$ a morphism $\varphi:\calF\to\calG$ is just a natural transformation.
\end{defn}
\begin{rmk}
	We say that $\varphi$ is injective (resp. surjective) if for all $x\in X$, 
	\[\varphi_x:\colim_{U\ni x}\calF(U)\to colim_{U\ni x}\calG(U)\]
	is injective (resp. surjective).
\end{rmk}

\begin{prop}
	The category of abelian sheaves form an Abelian category. :)
\end{prop}

\subsection{Operations with Sheaves}
Notice first that every sheaf is a presheaf. Then for any presheaf $\calF$, we can associate to it a sheaf 
by ``sheafifying.'' You can peek at Hartshorne for the actual definiton, but basically you add in the global sections that are missing (where local sections glue together to form them)
and remove global sections that are locally trivial.

More concretely, we get a functor:
\[()^+:\mathbf{Presh}_X\to\mathbf{Sh}_X\]
such that $()^+$ is left adjoint to the forgetful functor.

\begin{defn}
	The \textbf{global sections functor} is a functor 
	\[\Gamma(X,-):\mathbf{Sh}_X\to\Ab\]
	defined to be $\Gamma(X,\calF)=\calF(X)$.
\end{defn}
\begin{prop}
	$\Gamma(X,-)$ is left but not right exact.
\end{prop}
\begin{prf}
	That it is left exact can be proved a few ways, but Isaac mentioned that limits commute with limits which seems to be pretty slick.

	The primary counter-example that it is not right exact, consider the space $X=\bbC\setminus \{0\}$ and the sequence:
	\[0\to\underline\bbZ\hookrightarrow \calO\to \calO^\ast\to 0\]
	where $\underline\bbZ$ is the constant sheaf, $\calO$ is the holomorphic functions on $X$ and $\calO^\ast$ is the nonzero holomorphic functions on $X$. But then letting 
	the surjection be $f\mapsto e^{2\pi i f}$, there is clearly no global logarithm function in all cases. 
\end{prf}

\begin{thm}[Grothendieck `57]
	$\mathbf{Sh}_X$ has enough injectives.
\end{thm}

\subsection{Sites}
The motivation here is that there is a correspondence between affine schemes and commutative rings, and every scheme is glued together in such a way.
The problem here, though, is that he standard (Zariski) topology is just too coarse to glean anything useful.

The goal here is to put a topology on the category of schemes, forgetting completely about what the underlying topological space is. This motivates \'Etale coverings.

\begin{defn}
	Let $\calC$ be any category. A \textbf{Grothendieck topology} is a collection of distinguished collections indexed by objects in $\calC$ (think a collection of all coverings of a topological space)
	such that 
	\begin{itemize}
		\item Isomorphisms are coverings
		\item Coverings of coverings are coverings 
		\item Base changes of coverings are coverings: If $\{U_i\to X\}$ is a covering and if $V\to X$ is any morphism,
		then the fiber products (pullbacks) $\{U_i\times_XV\to V\}$ exist and are a covering.
	\end{itemize}
	Then $\calC$ along with its Grothendieck topology is called a \textbf{site.}
\end{defn}
\begin{ex}
	If $X$ is a topological space, $\mathbf{Op}(X)$ is a site, called the \textbf{small site} of $X$.

	$\mathbf{Top}$ is a site with jointly surjective open embeddings. Then $\mathbf{Top}/X$ is the \textbf{big site} of $X$.
\end{ex}

We're not going to talk about schemes here, but we will talk more about schemes.

\begin{defn}
	A \textbf{sheaf on a site} is a functor $\calF:\calC^{op}\to\mathbf{Set}$ satisfying the equalizer 
	diagram I wrote in definition~\ref{defn-sheaf-2}.

	An \textbf{abelian sheaf} is a sheaf that factors as $\calF:\calC^{op}\to\Ab$.
\end{defn}
\begin{ex}
	$\Hom_{\mathbf{Top}}(-,X):\mathbf{Top}^{op}\to\mathbf{Set}$. That this is a sheaf (on the big site of $X$) basically follows from the fact that 
	we have the gluing property in topological spaces.
\end{ex}
\begin{rmk}
	Notice that on the category of topological spaces, this tells us that all representable functors are sheaves!
\end{rmk}
\begin{ex}
	If $X$ is an abelian group object in $\mathbf{Top}$, then $\Hom(-,X)$ is an abelian functor.
\end{ex}
\begin{defn}
	$\Ab(\calC)$ is the category of abelian sheaves on $\calC$. 
\end{defn}
\begin{thm}
	$\Ab(\calC)$ is an abelian category with enough injectives.
\end{thm}
\begin{prop}
	For any object $X\in \calC$, we can define 
	\[\Gamma(X,-):\Ab(\calC)\to \Ab.\]

	Furthermore $\Gamma(X,-)$ is left exact, but not right exact.
\end{prop}

Then the goal is to measure failure of exactness of $\Gamma(T,-)$ where $T$ is a terminal object. Since we are in an abelian category we can do this in the usual way.

\begin{defn}
	If $\calF\in\Ab(\calC)$, $V\in \calC$, then the $i^{th}$ cohomology group of $\calF$ over $V$ is 
	\[H^i(V,\calF):=R^i\Gamma(V,\calI^\bullet)\]
	where $\calI^\bullet$ is an injective resolution of your $\calF$.
\end{defn}

@@

\newpage
\section{HW2 -- Due May}
\begin{prob}
	Let $R$ be a commutative ring, and $N$ be an $R$-module. Show that the following are equivalent:
	\begin{enumerate}
		\item $N$ is flat 
		\item $\Tor_i^R(M,N) = 0$ for any $i>0$ and any $R$-module $M$
		\item $\Tor_1^R(M,N) = 0$ for any $R$-module $M$
		\item $\Tor_1^R(R/I, N) = 0$ for any ideal $I \subset R$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Rightarrow$(b)$\Rightarrow$(c) is clear since $\Tor_i^R(-,N)$ is a left derived functor of $-\otimes N$. But when $N$ is flat, $-\otimes N$ is exact.

@@

\begin{prob}
	Let $(R, \frakm, k)$ be a Noetherian commutative local ring and $M$ be a finitely generated $R$-module. Prove that the following are equivalent: 
	\begin{enumerate}
		\item $M$ is free
		\item $M$ is flat
		\item The map $\frakm \otimes_R M \to R \otimes_R M =M$ induced by the embedding $\frakm \subset R$ is injective
		\item $\Tor_1(k,M)=0$.
	\end{enumerate}
\end{prob}
\begin{sol}
	(a)$\Rightarrow$(b): First let $M$ be free (and finitely generated). Then $M\cong R^k$ for some finite $k$. 
	Then take an injection $\iota:A\hookrightarrow B$ and consider the induced map $A\otimes_R M\to B\otimes_R M$. Since $M\cong R^k$, we get 
	that $A\otimes M\cong A^k$ and $B\otimes M\cong B^k$ with the induced map between them being $\iota$ on each summand. This is clearly injective.

	(b)$\Rightarrow$(c) follows since if $M$ is flat, $-\otimes_RM$ is exact, so it preserves injections. Thus in particular 
	the image of $\frakm\hookrightarrow R$ is also an injection.

	(c)$\Rightarrow$(d) Consider the short exact sequence 
	\[0\to\frakm\to R\to k\to 0\]
	and apply the long exact sequence in $\Tor$. But since the induced map from (c) is still injective, the following is exact:
	\[0=\Tor_1(R,M)\to \Tor_1(k,M)\to \frakm\otimes M\to R\otimes M\to k\otimes M\to 0\]
	where the equality on the left is because free modules are flat (see above).
	But then by the assumption in (c), this implies that $\Tor_1(k,M)$ injects into the trivial module, so it must be 0 as well.

	(d)$\Rightarrow$(a) Let $\{m_1,\dots,m_n\}$ be a minimal set of generators for $M$. Then define the map $f:R^n\to M$ via 
	$f(r_1,\dots,r_n)=r_1m_1+\cdots+r_nm_n.$ Let $I=\ker f$. Then we have the short exact sequence 
	\[0\to I\to R^n\to M\to 0\]
	and taking the long exact sequence associated to $-\otimes k$, we get 
	\[\cdots \Tor_1(k,M)\to I\otimes k\to R^n\otimes k\to M\otimes k\to 0\]
	and since $\Tor_1(k,M)=0$ and leveraging a couple more isomorphisms,
	\[0\to I/\frakm I\to k^n\to M/\frakm M\to 0\]
	is exact.

	But now by Nakayama, a basis for $M/\frakm M$ lifts to a minimal set of generators for $M$, but then $M/\frakm M\cong k^n$. Then by a dimension counting 
	argument, this forces $I/\frakm I=0.$ Since we assume that $R$ was Noetherian, $I$ is finitely generated, so again we can leverage Nakayama: since the quotient is 
	trivial, so is $I$ itself. But then $M\cong R^n$ and we're done. \textit{Thanks to Cody for giving me a hint to use Nakayama here.}
\end{sol}

@@

\begin{prob}
	Let $k$ be a field of positive characteristic $p$. 
	\begin{enumerate} 
		\item Show that the group algebra $kG$ for a finite group $G$ is Frobenius (in fact, symmetric). 
		\item Show that the restricted enveloping algebra $\fraku(\gl_n)$ is Frobenius. 
	\end{enumerate} 
\end{prob}
\begin{sol}
	\textbf{(a):} Define the form 
	\[\omega\left(\sum_{g\in G}a_gg,\sum_{g\in G}b_g g\right)=\sum_{g\in G}a_gb_g.\]

	One can easily check that this is a (symmetric!) nondegenerate form on $R\otimes R$. Note that I came up with this 
	by just guessing, but my first thought was to mirror the natural pairing we get with a vector space and its dual (linear functionals). Using this,
	one can construct (essentially) the same form. Perhaps this is what was intended, although my guesswork paid off this time. :)
\end{sol}
